---
title: "Get data for sibling paper"
output: html_notebook
---

This file uses `rdhs` and the DHS API to

* get some aggregate characteristics of DHS surveys
* download the Malawi 2000 DHS data, which we'll use later

## Get some aggregate characteristics of DHS surveys

```{r}
library(tidyverse)
library(siblingsurvival)
library(rdhs)
library(tictoc)
library(lubridate)
library(glue)
library(here)
library(wpp2017)

## these are optional - used to make a map
library(sf)
library(leaflet)
library(mapview)
```


Following the [rdhs vignette](https://cran.r-project.org/web/packages/rdhs/vignettes/introduction.html):

```{r}
set_rdhs_config(email = 'feehan@berkeley.edu', # you'll need your own account
                global=FALSE,
                project = "Sibling history example", # you'll need your own DHS-approved project
                cache_path = "./rdhs_cache",
                config_path = "rdhs.json"
                )
```

According to [the DHS API reference](https://api.dhsprogram.com/rest/dhs/surveycharacteristics/DHS), the maternal mortality module is survey characteristic 1.


```{r}
sc <- dhs_survey_characteristics()

# maternal mortality is SurveyCharacteristicID 1

ids <- dhs_countries(returnFields=c("CountryName", "DHS_CountryCode"))

survs <- dhs_surveys(surveyCharacteristicIds = 1) %>% left_join(dhs_countries())

survs.agg <- survs %>%
  rename(A3 = ISO3_CountryCode) %>%
  group_by(A3) %>%
  summarize(num_siblinghistory = n())

```

Some numbers - how many surveys, what time range, how many countries

```{r}
survs <- survs %>% mutate(year = year(ymd(FieldworkEnd)))

num.survs <- nrow(survs)
num.countries <- length(unique(survs$DHS_CountryCode))
mean.n <- mean(survs$NumberOfWomen)
min.year <- min(survs$year)
max.year <- max(survs$year)

glue("
    There are:
        {num.survs} surveys in
        {num.countries} countries
    that have included the sibling history module.

    Sibling history data have been collected from {min.year} to {max.year}.

    The average number of women interviewed is {format(round(mean.n,0),big.mark=',')}.

")
```


Look at the distribution of sample sizes for the Women's Questionnaire

```{r}

ggplot(survs) + 
  geom_histogram(aes(x=NumberOfWomen)) + 
  geom_vline(aes(xintercept=mean.n), color='red') +
  ggtitle(glue::glue("
Number of women interviewed
in DHS surveys with sibling history module
(Mean: {format(round(mean.n,0),big.mark=',')})
")) +
  theme_minimal()
```

```{r}
num.surveys <- nrow(survs)
num.countries <- length(unique(survs$ISO3_CountryCode))

glue::glue("The DHS sibling history module has been collected in {num.surveys} surveys in {num.countries} countries.")
```


Plot the availability of data by subregion

```{r}
survs.region <- survs %>%
  mutate(SubregionName = str_replace_all(SubregionName, " ", "\n"),
        subregion = fct_infreq(SubregionName)) %>%
  group_by(subregion) %>%
  summarize(tot = n()) 

sh.region <- ggplot(survs.region, aes(x=subregion, y=tot)) +
  geom_col() +
  xlab("") +
  ylab("Number of surveys\nwith sibling history data") +
  theme_minimal(base_size=14) +
  theme(text=element_text(size=20))

ggsave(filename=here('out', 'sh_byregion.pdf'),
       height=6, width=12,
       plot=sh.region)

sh.region
```

Plot the availability of data by year

```{r}
survs.year <- survs %>%
  mutate(year = str_sub(ReleaseDate, 1, 4)) %>%
  mutate(year_gp = cut(as.numeric(year), 
                       breaks=seq(from=1985, to=2020, by=5),
                       labels=paste(c("1985-\n1989", "1990-\n1994", "1995-\n1999", "2000-\n2004",
                                       "2005-\n2009", "2010-\n2014", "2015-\n2020"))))  %>%
  filter(! is.na(year_gp)) %>%
  group_by(year_gp) %>%
  summarize(tot = n()) 

sh.year <- ggplot(survs.year, aes(x=year_gp, y=tot)) +
  geom_col() +
  #xlab("Sub-region") +
  xlab("") +
  ylab("Number of surveys\nwith sibling history data") +
  theme_minimal(base_size=16) +
  theme(text=element_text(size=22))

ggsave(filename=here('out', 'sh_byyear.pdf'),
       height=6, width=9,
       plot=sh.year)

sh.year
```

Make a map

This geojson file downloaded from [geo-maps](https://github.com/simonepri/geo-maps/blob/master/info/countries-land.md):

```{r}
gjfn <- "countries-land-5km.geo.json"
countries <- sf::st_read(here('data', gjfn))
```

```{r}
c2 <- countries %>% 
  left_join(survs.agg) %>%
  mutate(num_siblinghistory = ifelse(is.na(num_siblinghistory),
                                     0,
                                     num_siblinghistory))
```

```{r}
#pal <- colorNumeric("viridis", NULL)
pal <- colorBin("YlOrRd", domain = c2$num_siblinghistory)

m <- leaflet(c2) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
    fillColor = ~pal(num_siblinghistory),
    label = ~paste0(A3, ": ", formatC(num_siblinghistory, big.mark = ","))) %>%
  setView(0,0, zoom=2) %>%
  addLegend(pal = pal, values = ~num_siblinghistory, opacity = 1.0,
            position="bottomright")

mapshot(m, file=here('out', "sh_count.png"))

m
```


## Get the data from the Malawi 2000 DHS

Using the `rdhs` package, download the Malawi 2000 survey.

We want the Malawi 2000 DHS

```{r}
svy_ids <- dhs_surveys(surveyYear=c(2000), 
                       surveyType="DHS",
                       countryIds=c("MW") # the country code for Malawi
                       )
```

```{r}
mw2000_fn <- dhs_datasets(surveyIds = 'MW2000DHS',
                          fileFormat = "FL")
mw2000_ir_fn <- mw2000_fn %>% filter(FileType=='Individual Recode') %>% pull(FileName)
```

Download the actual dataset

```{r}
downloads <- get_datasets(dataset_filenames=c(mw2000_ir_fn))
```

```{r}
mw2000 <- readRDS(downloads$MWIR41FL)
```

# prep the sibling histories

```{r}
prepped <- prep_dhs_sib_histories(mw2000,
                                  # NB: default varmap is sibhist_varmap_dhs6
                                  keep_missing = FALSE,
                                  keep_varmap_only = TRUE)

mw.ego <- prepped$ego.dat
mw.sib <- prepped$sib.dat
```

Here's the respondent (ego) data

```{r}
mw.ego
```

And here's the sibling data

```{r}
mw.sib
```

Let's add an indicator for whether or not each sib is in the frame population

```{r}
mw.sib <- mw.sib %>% 
  mutate(in.F = as.numeric((sib.alive==1) & (sib.age >= 15) & (sib.age <= 49) & (sib.sex == 'f')))
```

```{r}
with(mw.sib, table(in.F, useNA='ifany'))
```

```{r}
mw.sib <- mw.sib %>% filter(! is.na(in.F)) 
```

For the simulation study, we'll need a weight that is scaled to the size of the frame population.

```{r}
## get Malawi estimated popn in 2000 (the survey year)
## this comes from the wpp2017 package, i.e., UNPD estimates
data(popF)
malawi.N.F <- popF %>%
  filter(name=='Malawi') %>%
  filter(age %in% c('15-19', '20-24', '25-29', '30-34', '35-39',
                    '40-45', '45-49', '50-54', '55-59', '60-64')) %>%
  pull(`2000`) %>%
  sum()
malawi.N.F <- malawi.N.F * 1e3

# scale factor to convert relative to absolute weights
malawi.sf <- malawi.N.F / sum(mw.ego$wwgt)

mw.ego <- mw.ego %>%
  mutate(scaledweight = wwgt * malawi.sf)

mw.sib <- mw.sib %>%
  mutate(scaledweight = wwgt * malawi.sf)
```


```{r}
save(mw.ego,
     mw.sib,
     file=here('data', 'malawi_dhs_2000.RData'))
```






