---
title: "Malawi 2000 DHS sibling history analysis"
output: html_notebook
---

## Analyze the Malawi 2000 sibling histories as an illustration of our methods

```{r}
library(tidyverse)
library(siblingsurvival)
library(rdhs)
library(tictoc)
library(lubridate)
library(glue)
library(here)
library(latex2exp)
library(kableExtra)

tic("Malawi analysis")
```

Load the data prepped in `01_get_data.Rmd`

```{r}
load(file=here('data', 'malawi_dhs_2000.RData'))
```

Create an object that describes the cells we'd like to use for the estimates

```{r}
cc <- cell_config(age.groups='5yr_to50', 
                  time.periods='7yr_beforeinterview',
                  start.obs='sib.dob',    # date of birth
                  end.obs='sib.endobs',   # either the date respondent was interviewed (if sib is alive) or date of death (if sib is dead)
                  event='sib.death.date', # date of death (for sibs who died)
                  age.offset='sib.dob',   # date of birth
                  time.offset='doi',      # date of interview
                  exp.scale=1/12)
```


### Estimating death rates

Given these preparatory steps, the `sibling_estimator` function will take care of estimating
death rates from the sibling histories for us.

```{r}
ex_ests <- sibling_estimator(sib.dat = mw.sib,
                             ego.id = 'caseid',            # column with the respondent id
                             sib.id = 'sibid',             # column with sibling id 
                                                           # (unique for each reported sibling)
                             sib.frame.indicator = 'in.F', # indicator for sibling frame population membership
                             sib.sex = 'sib.sex',          # column with sibling's sex
                             cell.config=cc,               # cell configuration we created above
                             weights='wwgt')               # column with the respondents' sampling weights

names(ex_ests)
```

`sibling_estimator` returns a list with the results. We'll focus on `asdr.ind`, which has the individual visibility
estimates, and `asdr.agg`, which has the aggregate visibility estimates.

Here are the individual visibility estimates:

```{r}
ex_ests$asdr.ind
```

And here are the aggregate visibility estimates

```{r}
ex_ests$asdr.agg
```

### Plotting the results

We'll make some plots showing the results


```{r}
fig.height <- 4
fig.width <- 5

ind.ests <- ggplot(ex_ests$asdr.ind) +
  geom_line(aes(x=sib.age, y=1000*asdr.hat, color=sib.sex, group=sib.sex)) +
  theme_minimal() +
  scale_y_log10() + 
  ggtitle('individual visibility estimator, 7yr before survey')

ggsave(filename=here('out', 'mw2000-ind-ests.pdf'),
       plot=ind.ests,
       height=fig.height, width=fig.width)

ind.ests
```


```{r}
fig.height <- 4
fig.width <- 5

agg.ests <- ggplot(ex_ests$asdr.agg) +
  geom_line(aes(x=sib.age, y=1000*asdr.hat, color=sib.sex, group=sib.sex)) +
  theme_minimal() +
  scale_y_log10() + 
  ggtitle('aggregate visibility estimator, 7yr before survey')

ggsave(filename=here('out', 'mw2000-agg-ests.pdf'),
       plot=agg.ests,
       height=fig.height, width=fig.width)

agg.ests
```

```{r}
fig.height <- 4
fig.width <- 8

compare <- bind_rows(ex_ests$asdr.ind, 
                     ex_ests$asdr.agg) %>%
  ungroup() %>%
  mutate(sib.sex = dplyr::recode(sib.sex,
                                 `m` = "Males",
                                 `f` = "Females"),
         sib.agelab = stringr::str_sub(sib.age,2,3),
         estimator = dplyr::recode(estimator,
                                   `sib_agg`='Aggregate visibility',
                                   `sib_ind`='Individual visibility'))


comp.plot <- ggplot(compare) +
  geom_line(aes(x=sib.agelab, 
                y=1000*asdr.hat, 
                color=estimator, 
                group=estimator)) +
  theme_minimal() +
  facet_grid(. ~ sib.sex) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_y_log10() + 
  ylab(TeX('$\\widehat{M}^V_{\\alpha} \\times 1,000$')) +
  xlab("") +
  theme(legend.position="bottom")
  #scale_y_log10() + 
  #ggtitle('both estimators, 7yr before survey')

ggsave(filename=here('out', 'mw2000-comp-ests.pdf'),
       plot=comp.plot,
       height=fig.height, width=fig.width)

comp.plot
```


## Variance estimates

In practice, we want point estimates and estimated sampling uncertainty for the death rates.
We'll use the rescaled bootstrap to estimate sampling uncertainty. This can be done with the `surveybootstrap` package.

Before we use the rescaled bootstrap, we need to know a little bit about the sampling design of the survey
we're working with.
In this example dataset, we have a stratified, multistage design. So we'll need to tell the bootstrap function about
the strata and the primary sampling units. In this example data, these are indicated by the 'stratum' and 'psu'
columns of the dataset.

(NOTE: this takes about a minute or so on a 2018 MBP for 1000 resamples.)

```{r}
set.seed(101010)

tic('running bootstrap')
## this will take a little while -- for 1000 reps, it takes about 1 minutes on a 2018 Macbook Pro
num.boot.reps <- 1000
bootweights <- surveybootstrap::rescaled.bootstrap.weights(survey.design = ~ psu + 
                                                                            stratum(stratum_analysis),
                                                           # a high number is good here, though that will obviously
                                                           # make everything take longer
                                                           num.reps=num.boot.reps,
                                                           idvar='caseid',       # column with the respondent ids
                                                           weights='wwgt', # column with the sampling weight
                                                           survey.data=mw.ego    # note that we pass in the respondent data, NOT the sibling data
                                                           )
toc()
```


The result, `bootweights`, is a dataframe that has a row for each survey respondent, a column for the survey respondent's ID,
and then `num.boot.reps` columns containing survey weights that result from the bootstrap procedure.
The basic idea is to calculate estimated death rates using each one of the `num.boot.reps` sets of weights.
The variation across the estimates is then an estimator for the sampling variation.

To make this easier, the `sibling_estimator` function can take a dataset with bootstrap resampled weights; it will
then calculate and summarize the estimates for you.

(NOTE: this takes about 3 minutes or so on a 2018 MBP when `bootweights` has 1000 resamples.)

```{r}
# to save time, we'll only use a subset of the bootstrap replicates

short.bootweights <- bootweights %>% select(1:11)
# toggle between short bootweights (faster, for coding) and long bootweights (for realism)
#ests.bootweights <- short.bootweights
ests.bootweights <- bootweights

tic('calculating estimates with bootstrap')
mw_boot_ests <- sibling_estimator(sib.dat = mw.sib,
                                  ego.id = 'caseid',
                                  sib.id = 'sibid',
                                  sib.frame.indicator = 'in.F',
                                  sib.sex = 'sib.sex',
                                  cell.config=cc,
                                  boot.weights=ests.bootweights,  # to get sampling uncertainty, we pass boot.weights into sibling_estimator
                                  return.boot=TRUE,                # when TRUE, return all of the resampled estimates (not just summaries)
                                  weights='wwgt')
toc()
```


Finally, let's plot the estimated death rates along with their sampling uncertainty:

```{r}
fig.height <- 4
fig.width <- 5

ind.ests.withci <- ggplot(mw_boot_ests$asdr.ind) +
  geom_ribbon(aes(x=sib.age, ymin=1000*asdr.hat.ci.low, ymax=1000*asdr.hat.ci.high, fill=sib.sex, group=sib.sex), alpha=.2) +
  geom_line(aes(x=sib.age, y=1000*asdr.hat, color=sib.sex, group=sib.sex)) +
  theme_minimal() +
  scale_y_log10() + 
  ggtitle('individual visibility estimator, 7yr before survey')

ggsave(filename=here('out', 'mw2000-ind-ests-withci.pdf'),
       plot=ind.ests.withci,
       height=fig.height, width=fig.width)

ind.ests.withci
```



```{r}
fig.height <- 4
fig.width <- 5

agg.ests.withci <- ggplot(mw_boot_ests$asdr.agg) +
  geom_ribbon(aes(x=sib.age, ymin=1000*asdr.hat.ci.low, ymax=1000*asdr.hat.ci.high, fill=sib.sex, group=sib.sex), alpha=.2) +
  geom_line(aes(x=sib.age, y=1000*asdr.hat, color=sib.sex, group=sib.sex)) +
  theme_minimal() +
  scale_y_log10() + 
  ggtitle('aggregate visibility estimator, 7yr before survey')

ggsave(filename=here('out', 'mw2000-agg-ests-withci.pdf'),
       plot=agg.ests.withci,
       height=fig.height, width=fig.width)

agg.ests.withci
```

Figure comparing male to male and female to female estimates

```{r}
fig.height <- 4
fig.width <- 8

comp.withci <- bind_rows(mw_boot_ests$asdr.agg,
                         mw_boot_ests$asdr.ind) %>%
  ungroup() %>%
  mutate(sib.sex = dplyr::recode(sib.sex,
                                 `m` = "Males",
                                 `f` = "Females"),
         sib.agelab = stringr::str_sub(sib.age,2,3),
         estimator = dplyr::recode(estimator,
                                   `sib_agg`='Aggregate visibility',
                                   `sib_ind`='Individual visibility'))

comp.ests.withci <- ggplot(comp.withci) +
  geom_ribbon(aes(x=sib.agelab, 
                  ymin=1000*asdr.hat.ci.low, 
                  ymax=1000*asdr.hat.ci.high, 
                  fill=estimator, 
                  group=estimator), 
              alpha=.2) +
  geom_line(aes(x=sib.agelab, 
                y=1000*asdr.hat, 
                color=estimator, 
                group=estimator)) +
  facet_grid(. ~ sib.sex) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_log10() + 
  ylab(TeX('$\\widehat{M}^V_{\\alpha} \\times 1,000$')) +
  xlab("") +
  theme(legend.position="bottom")
  #ggtitle('sibling history estimators, 7yr before survey')

ggsave(filename=here('out', 'mw2000-comp-ests-withci.pdf'),
       plot=comp.ests.withci,
       height=fig.height, width=fig.width)

comp.ests.withci
```

Calculate some summaries about sampling error across the two estimators

```{r}
comp.ci <- bind_rows(mw_boot_ests$asdr.ind, mw_boot_ests$asdr.agg)

comp.ci <- comp.ci %>%
  # calculate the relative standard error
  mutate(rel.se = asdr.hat.se / asdr.hat) %>%
  group_by(estimator) %>%
  # take the average
  summarize(mean.rel.se = mean(rel.se),
            gmean.rel.se = exp(mean(log(rel.se))))

comp.ci.fortab <-
  comp.ci %>%
  select(estimator, mean.rel.se) %>%
  mutate(estimator = dplyr::recode(estimator,
                                   `sib_agg`="Aggregate visibility",
                                   `sib_ind`="Individual visibility"))

comp.ci.tab <- knitr::kable(comp.ci.fortab,
             format='latex',
             booktabs=T,
             label="ci-compare",
             col.names=c("Estimator", "Estimated Avg. Relative Std. Error"),
             escape=FALSE,
             format.args=list(digits=2),
             caption="Comparison in average relative standard error for aggregate and individual visibility estimates of death rates across all ages and sexes.")

cat(comp.ci.tab, 
    file=here('out', "mw-compare-ci.tex"))
```

Now repeat the summaries, breaking them down by sex.

```{r}
comp.ci.bysex <- bind_rows(mw_boot_ests$asdr.ind, mw_boot_ests$asdr.agg)

comp.ci.bysex <- comp.ci.bysex %>%
  # calculate the relative standard error
  mutate(rel.se = asdr.hat.se / asdr.hat) %>%
  group_by(estimator, sib.sex) %>%
  # take the average
  summarize(mean.rel.se = mean(rel.se),
            gmean.rel.se = exp(mean(log(rel.se))))

comp.ci.bysex.fortab <-
  comp.ci.bysex %>%
  ungroup() %>%
  select(estimator, sib.sex, mean.rel.se) %>%
  spread(sib.sex, mean.rel.se)   %>%
  mutate(estimator = dplyr::recode(estimator,
                                   `sib_agg`="Aggregate visibility",
                                   `sib_ind`="Individual visibility")) %>%
  ## be sure the order is fixed, so columns are right
  select(estimator, f, m)

comp.ci.bysex.tab <- knitr::kable(comp.ci.bysex.fortab,
             format='latex',
             booktabs=T,
             label="ci-compare-bysex",
             col.names=kableExtra::linebreak(c("Estimator", "Females",  "Males")),
             escape=FALSE,
             format.args=list(digits=2),
             caption="Comparison in average relative standard error for aggregate and individual visibility estimates of death rates across all ages for males and for females.") %>%
             add_header_above(c("", "Estimated Average\nRelative Standard Error"=2))

cat(comp.ci.bysex.tab, 
    file=here('out', "mw-compare-ci-bysex.tex"))
```

## Internal consistency checks

In this section, we illustrate how to conduct internal consistency checks


```{r}
# to produce single year of age IC checks, we need to set
# age.groups to '1yr'
# time.periods to '12mo_beforeinterview'
cc_sya <- cell_config(age.groups='1yr_to50', 
                      time.periods='12mo_beforeinterview',
                      start.obs='sib.dob',    # date of birth
                      end.obs='sib.endobs',   # either the date respondent was interviewed (if sib is alive) or date of death (if sib is dead)
                      event='sib.death.date', # date of death (for sibs who died)
                      age.offset='sib.dob',   # date of birth
                      time.offset='doi',      # date of interview
                      exp.scale=1/12)
```


This next chunk takes longer to run, about 15 minutes on a 2018 MBP.

```{r}
short.bootweights <- bootweights %>% select(1:11)

# toggle between short bootweights (faster, for coding) and long bootweights (for realism)
#ests.bootweights <- short.bootweights
ests.bootweights <- bootweights

tic('calculating estimates with bootstrap')
mw_boot_sya_ests <- sibling_estimator(sib.dat = mw.sib,
                                      ego.id = 'caseid',
                                      sib.id = 'sibid',
                                      sib.frame.indicator = 'in.F',
                                      sib.sex = 'sib.sex',
                                      cell.config=cc_sya,
                                      boot.weights=ests.bootweights,  # to get sampling uncertainty, we pass boot.weights into sibling_estimator
                                      return.boot=TRUE,                # when TRUE, return all of the resampled estimates (not just summaries)
                                      weights='wwgt')
toc()
```

```{r}
mw_boot_sya_ests$esc.dat <- mw_boot_sya_ests$esc.dat %>%
  mutate(sib.sya = as.numeric(str_sub(sib.age, 2, 3)))

# toggle between short bootweights (faster, for coding) 
# and long bootweights (for realism)
#ic.sya.bootweights <- short.bootweights
ic.sya.bootweights <- bootweights

tic("Internal consistency checks - single year of age")
ic.checks.sya <- sib_ic_checks(mw_boot_sya_ests$esc.dat,
                               ego.dat=mw.ego,
                               ego.id='caseid',
                               sib.id='sibid',
                               sib.frame.indicator='in.F',
                               ## we're doing single year of age here
                               #sib.cell.vars=c('sib.age', 'sib.sex'),
                               sib.cell.vars=c('sib.sya', 'sib.sex'),
                               ego.cell.vars=c('age', 'sex'),
                               boot.weights=ic.sya.bootweights)
toc()

names(ic.checks.sya)
```


```{r}
fig.height <- 4
fig.width <- 7

ic.sya.plot <- ggplot(ic.checks.sya$ic.summ) +
  geom_hline(yintercept=0) +
  geom_pointrange(aes(x=age,
                      y=diff_mean,
                      ymin=diff_ci_low,
                      ymax=diff_ci_high)) +
  theme_minimal() +
  #ggtitle("Internal consistency checks") +
  xlab("") +
  ylab(expression(Delta[alpha]))

ggsave(filename=here('out', 'mw2000-ic-sya.pdf'),
       plot=ic.sya.plot,
       height=fig.height, width=fig.width)

ic.sya.plot
```

```{r}
fig.height <- 4
fig.width <- 5

ic.sya.plot.smooth <- ggplot(ic.checks.sya$ic.summ %>% 
                               mutate(age = as.numeric(age))) +
  geom_hline(yintercept=0) +
  geom_point(aes(x=age,
                 y=diff_mean)) +
  geom_pointrange(aes(x=age,
                      y=diff_mean,
                      ymin=diff_ci_low,
                      ymax=diff_ci_high)) +
  geom_smooth(aes(x=age, y=diff_mean), color='blue') +
  theme_minimal() +
  ggtitle("Internal consistency checks") +
  xlab("Age") +
  ylab(expression(Delta[alpha]))

ggsave(filename=here('out', 'mw2000-ic-sya-withsmooth.pdf'),
       plot=ic.sya.plot.smooth,
       height=fig.height, width=fig.width)

ic.sya.plot.smooth
```


## Visibilities

```{r}
sib.F.dat <- mw.sib %>%
  group_by(caseid) %>%
  summarize(y.F = sum(in.F),
            sib.size = n() + 1)

ego.vis <- mw.ego %>% 
  select(caseid, wwgt, age.cat, sex) %>%
  left_join(sib.F.dat, by='caseid')

# if nothing was joined in, there are no sibs
ego.vis <- ego.vis %>%
  mutate(y.F=ifelse(is.na(y.F), 0, y.F),
         y.Fplusone = y.F + 1,
         sib.size = ifelse(is.na(sib.size), 1, sib.size))

# weighted harmonic mean
wh.mean <- function(x, w) {
  return(sum(w) / sum(w/x))
}

S.hat <- wh.mean(ego.vis$y.Fplusone, ego.vis$wwgt)
S.adj.factor <- 1 - (1/S.hat)
  
y.F.bar <- weighted.mean(ego.vis$y.F, ego.vis$wwgt)
approx.S.hat <- y.F.bar + 1
approx.S.adj.factor <- 1 - (1/approx.S.hat)

ego.vis.agg <- ego.vis %>%
  mutate(age.cat = paste(age.cat)) %>%
  group_by(sex, age.cat) %>%
  summarise(y.F.bar = weighted.mean(y.F, wwgt)) %>%
  mutate(adj.factor = S.adj.factor,
         # this is the all-ages approximation
         adj.factor.approx1 = approx.S.adj.factor,
         # this is the age-specific approximation
         adj.factor.approx2 = y.F.bar / (y.F.bar + 1))

```

Plot fraction of egos with zero visibility

```{r}
ego.vis.toplot <-
  ego.vis %>%
  mutate(is_invisible = as.numeric(y.F == 0)) %>%
  group_by(sex, age.cat) %>%
  summarize(frac_invisible = sum(is_invisible * wwgt) / sum(wwgt))
```

```{r}
fig.height <- 4
fig.width <- 4

ego.invis.plot <- ggplot(ego.vis.toplot %>%
                           mutate(age.lab = str_sub(age.cat, 2, 3))) +
  geom_point(aes(x = age.lab,
                 y = frac_invisible)) +
  expand_limits(y=0) +
  scale_y_continuous(labels=scales::percent) +
  theme_minimal() +
  xlab("Age group") +
  ylab("Estimated % of respondents\nwith zero visibility\n(Malawi 2000 DHS)")

ggsave(filename=here('out', 'mw2000-invis-est.pdf'),
       plot=ego.invis.plot,
       height=fig.height, width=fig.width)

ego.invis.plot
```


Make adjusted individual estimates

```{r}
for.adj.agg <- ex_ests$asdr.agg %>%
  left_join(ego.vis.agg,
            by=c('sib.sex'='sex', 'sib.age'='age.cat')) %>%
  mutate(adj.factor = ifelse(is.na(adj.factor), 1, adj.factor),
         adj.factor.approx1 = ifelse(is.na(adj.factor.approx1), 1, adj.factor.approx1),
         adj.factor.approx2 = ifelse(is.na(adj.factor.approx2), 1, adj.factor.approx2))

adj.agg.ests <- for.adj.agg %>% 
  mutate(estimator = 'sib_agg_nocovar_adj',
         asdr.hat = asdr.hat * adj.factor)

adj.agg.ests.approx1 <- for.adj.agg %>% 
  mutate(estimator = 'sib_agg_approx1_adj',
         asdr.hat = asdr.hat * adj.factor.approx1)

adj.agg.ests.approx2 <- for.adj.agg %>% 
  mutate(estimator = 'sib_agg_approx2_adj',
         asdr.hat = asdr.hat * adj.factor.approx2)

```

And plot a comparison (for the appendix, with all versions)

```{r}
fig.height <- 6
fig.width <- 7

compare <- bind_rows(ex_ests$asdr.ind, 
                     ex_ests$asdr.agg, 
                     #adj.agg.ests,
                     adj.agg.ests.approx1,
                     adj.agg.ests.approx2) %>%
  ungroup() %>%
  mutate(sib.sex = dplyr::recode(sib.sex,
                                 `m` = "Males",
                                 `f` = "Females"),
         sib.agelab = stringr::str_sub(sib.age,2,3),
         estimator = dplyr::recode(estimator,
                                   `sib_agg`='Aggregate',
                                   #`sib_agg_nocovar_adj`='Aggregate, adjusted (no covariance)',
                                   `sib_agg_approx1_adj`='Aggregate, adjusted\n(all age approximation)',
                                   `sib_agg_approx2_adj`='Aggregate, adjusted\n(age-specific approximation)',
                                   `sib_ind`='Individual'))

compare.plot <- ggplot(compare) +
  geom_line(aes(x=sib.agelab, 
                y=1000*asdr.hat, 
                #color=sib.sex, 
                #color=adjusted, 
                color=estimator, 
                group=interaction(estimator, sib.sex))) +
  theme_minimal() +
  #facet_grid(sex ~ .) +
  xlab("") +
  ylab(TeX('$\\widehat{M}^V_{\\alpha} \\times 1,000$')) +
  facet_grid(. ~ sib.sex) +
  #scale_linetype_manual(values=c('Aggregate'='solid',
  #                               'Aggregate, adjusted'='dashed',
  #                               'Individual'='dotted'),
  #                      name='Estimator') +
  scale_y_log10() +
  guides(color=guide_legend(direction='vertical')) +
  theme(legend.position="bottom")
  #ggtitle('sibling estimators, 7yr before survey')

ggsave(filename=here('out', 'mw2000-adj-agg-vis-all.pdf'),
       plot=compare.plot,
       height=fig.height, width=fig.width)

compare.plot
```

And also plot a comparison with only one version, for the main paper

```{r}
fig.height <- 4
fig.width <- 7

compare <- bind_rows(ex_ests$asdr.ind, 
                     ex_ests$asdr.agg, 
                     #adj.agg.ests) %>%
                     adj.agg.ests.approx1) %>%
                     #adj.agg.ests.approx2) %>%
  ungroup() %>%
  mutate(sib.sex = dplyr::recode(sib.sex,
                                 `m` = "Males",
                                 `f` = "Females"),
         sib.agelab = stringr::str_sub(sib.age,2,3),
         estimator = dplyr::recode(estimator,
                                   `sib_agg`='Aggregate',
                                   #`sib_agg_nocovar_adj`='Aggregate, adjusted (no covariance)',
                                   `sib_agg_approx1_adj`='Aggregate, adjusted\n(approximation)',
                                   #`sib_agg_approx2_adj`='Aggregate, adjusted (age-specific approximation)',
                                   `sib_ind`='Individual'))

compare.plot <- ggplot(compare) +
  geom_line(aes(x=sib.agelab, 
                y=1000*asdr.hat, 
                #color=sib.sex, 
                #color=adjusted, 
                linetype=estimator, 
                group=interaction(estimator, sib.sex))) +
  theme_minimal() +
  #facet_grid(sex ~ .) +
  xlab("") +
  ylab(TeX('$\\widehat{M}^V_{\\alpha} \\times 1,000$')) +
  facet_grid(. ~ sib.sex) +
  scale_linetype_manual(values=c('Aggregate'='solid',
                                 #'Aggregate, adjusted (no covariance)'='dashed',
                                 'Aggregate, adjusted\n(approximation)'='dashed',
                                 #'Aggregate, adjusted (age-specific approximation)'='dashed',
                                 'Individual'='dotted'),
                        name='Estimator:') +
  scale_y_log10() +
  #guides(linetype=guide_legend(direction='vertical')) +
  theme(legend.position="bottom")
  #ggtitle('sibling estimators, 7yr before survey')

ggsave(filename=here('out', 'mw2000-adj-agg-vis.pdf'),
       plot=compare.plot,
       height=fig.height, width=fig.width)

compare.plot
```

## Estimates 1 year before survey

```{r}
cc_12mo <- cell_config(age.groups='5yr_to50',
                       time.periods='12mo_beforeinterview',
                       #time.periods='7yr_beforeinterview',
                       start.obs='sib.dob',    # date of birth
                       end.obs='sib.endobs',   # either the date respondent was interviewed 
                                               # (if sib is alive) or date of death (if sib is dead)
                       event='sib.death.date', # date of death (for sibs who died)
                       age.offset='sib.dob',   # date of birth
                       time.offset='doi',      # date of interview
                       exp.scale=1/12)
```


Given these preparatory steps, the `sibling_estimator` function will take care of estimating
death rates from the sibling histories for us.

```{r}
tic('Estimates 1 yr before survey')
mw_ests_12mo <- sibling_estimator(sib.dat = mw.sib,
                                  ego.id = 'caseid',            # column with the respondent id
                                  sib.id = 'sibid',             # column with sibling id 
                                                                # (unique for each reported sibling)
                                  sib.frame.indicator = 'in.F', # indicator for sibling frame population membership
                                  sib.sex = 'sib.sex',          # column with sibling's sex
                                  cell.config=cc_12mo,          # cell configuration we created above
                                  boot.weights=ests.bootweights,# to get sampling uncertainty, we pass boot.weights into sibling_estimator
                                  return.boot=TRUE,             # when TRUE, return all of the resampled estimates (not just summaries)
                                  weights='wwgt')               # column with the respondents' sampling weights
toc()

names(mw_ests_12mo)
```

```{r}
fig.height <- 4
fig.width <- 8

comp.withci.12mo <- bind_rows(mw_ests_12mo$asdr.agg,
                              mw_ests_12mo$asdr.ind) %>%
  ungroup() %>%
  mutate(sib.sex = dplyr::recode(sib.sex,
                                 `m` = "Males",
                                 `f` = "Females"),
         sib.agelab = stringr::str_sub(sib.age,2,3),
         estimator = dplyr::recode(estimator,
                                   `sib_agg`='Aggregate visibility',
                                   `sib_ind`='Individual visibility'))

comp.withci.12mo <- ggplot(comp.withci.12mo) +
  geom_ribbon(aes(x=sib.agelab, 
                  ymin=1000*asdr.hat.ci.low, 
                  ymax=1000*asdr.hat.ci.high, 
                  fill=estimator, 
                  group=estimator), 
              alpha=.2) +
  geom_line(aes(x=sib.agelab, 
                y=1000*asdr.hat, 
                color=estimator, 
                group=estimator)) +
  facet_grid(. ~ sib.sex) +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_log10() + 
  ylab(TeX('$\\widehat{M}^V_{\\alpha} \\times 1,000$')) +
  xlab("") +
  theme(legend.position="bottom")
  #ggtitle('sibling history estimators, 7yr before survey')

ggsave(filename=here('out', 'mw2000-comp-ests-12mo.pdf'),
       plot=comp.withci.12mo,
       height=fig.height, width=fig.width)

comp.withci.12mo
```

## Sensitivity to invisible deaths

**Harmonic mean**  

Relative error in using visible death rate to estimate total death rate is

$$
\text{RE} = p^I_{D_\alpha} \left(\frac{1-K}{K}\right)
$$

```{r}
# grid size
m_grid <- 100

re_pD <- function(p_i_D, K) {
  return( p_i_D * ((1-K)/K))
}
  
input <- list(scale_K_D = .2,
              scale_p_D = .2,
              p_i_D   = .2,
              K       = 1)

#p.vals <- seq(from=(1-input$scale_p_D)*input$p_i_D, 
#              to=  (1+input$scale_p_D)*input$p_i_D, 
#              length.out=m_grid)
p.vals <- seq(from=.15, 
              to=  .30, 
              length.out=m_grid)

K.vals <- seq(from=(1-input$scale_K_D)*input$K, 
              to=  (1+input$scale_K_D)*input$K, 
              length.out=m_grid)

toplot.D <- expand.grid(p_i_D = p.vals,
                        K = K.vals)

toplot.D$m_v_relerr <- map2_dbl(toplot.D$p_i_D,
                                toplot.D$K, 
                                ~ re_pD(p_i_D=.x, K=.y))
```

```{r}
fig.height <- 4
fig.width <- 4

invis.sens.pD <- ggplot(toplot.D, 
       aes(x=p_i_D, y=K, z = m_v_relerr)) +
       geom_raster(aes(fill=m_v_relerr)) +
       geom_contour(bindwidth=.01, color='white', direction='horizontal') +
       theme_minimal() +
       xlab(TeX("p^I_{D}")) +
       scale_fill_gradient2(labels=scales::percent) +
       theme(legend.position='bottom') +
       guides(fill=guide_colorbar(title="Relative\nerror",
                                  barwidth=15)) 

ggsave(filename=here('out', 'invis-sens-pD.pdf'),
       plot=invis.sens.pD,
       height=fig.height, width=fig.width)

invis.sens.pD
```

**Arithmetic mean**  

Relative error in using visible death rate to estimate total death rate is

$$
\text{RE} = \frac{p^I_{N_\alpha} (1-K)}{1 - p^I_{N_\alpha}(1-K)}
$$

```{r}
# grid size
m_grid <- 100

re_pN <- function(p_i_N, K) {
  return( (p_i_N * (1-K)) / (1 - (p_i_N * (1-K))))
}
  
input <- list(scale_K_N = .2,
              scale_p_N = .5,
              p_i_N   = .225,
              K       = 1)

p.vals <- seq(from=.15, 
              to=  .30, 
              length.out=m_grid)
#p.vals <- seq(from=(1-input$scale_p_N)*input$p_i_N, 
#              to=  (1+input$scale_p_N)*input$p_i_N, 
#              length.out=m_grid)

K.vals <- seq(from=(1-input$scale_K_N)*input$K, 
              to=  (1+input$scale_K_N)*input$K, 
              length.out=m_grid)

toplot.N <- expand.grid(p_i_N = p.vals,
                        K = K.vals)

toplot.N$m_v_relerr <- map2_dbl(toplot.N$p_i_N,
                                toplot.N$K, 
                                ~ re_pN(p_i_N=.x, K=.y))
```

```{r}
fig.height <- 4
fig.width <- 4

invis.sens.pN <- ggplot(toplot.N, 
       aes(x=p_i_N, y=K, z = m_v_relerr)) +
       geom_raster(aes(fill=m_v_relerr)) +
       geom_contour(bindwidth=.01, color='white', direction='horizontal') +
       theme_minimal() +
       xlab(TeX("p^I_{N}")) +
       scale_fill_gradient2(labels=scales::percent) +
       theme(legend.position='bottom') +
       guides(fill=guide_colorbar(title="Relative\nerror",
                                  barwidth=15)) 

ggsave(filename=here('out', 'invis-sens-pN.pdf'),
       plot=invis.sens.pN,
       height=fig.height, width=fig.width)

invis.sens.pN
```

```{r}
toc()
```

