---
title: "Analyze the results of the DHS sibling history simulations"
output: html_notebook
---

```{r}
library(surveybootstrap)
library(networkreporting)
library(dplyr)
library(tidyverse)
library(forcats)
library(haven)
library(igraph)
library(lazyeval)
library(nrsimulatr)

library(tictoc)
library(ggplot2)
library(latex2exp)
library(glue)

library(here)

tic("Analyzing simulation results")
```

Load the results of the simulation

```{r}
load(here('sim-data', 'sib-simulated-surveys.RData'))
```

NOTE: we'll only look at age groups 2 to 8, i.e., ages 15 to 50

```{r}
svy.sims.forplots <- 
  svy.sims %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2'))

all.truths.forplots <-
  all.truths %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2'))
  
all.estimands.forplots <-
  all.estimands %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2'))

adj.factors.agg.forplots <-
  adj.factors.agg %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2'))

adj.factors.ind.forplots <-
  adj.factors.ind %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2'))

visible_truth.forplots <-
  visible_truth %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2'))

agegp.labels.used <- agegp.labels[-c(1,9)]
```


## Plot the (simulated) underlying true death rates

```{r}
qty.names <- c('Total population death rate'='truth',
               'Visible population death rate'='visible_truth',
               'Visible population (inc. resp.) death rate' = 'visible_truth_withr',
               'Invisible population death rate'='invisible_truth')

sex.labs <- c('Females'='f',
              'Males'='m')

all.truth.plot <- all.truths.forplots %>%
  separate(cell, into=c('agegp', 'sex'), sep='X') %>%
  mutate(agegp=str_extract(agegp, '\\d'),
         sex=str_extract(sex, '\\d')) %>%
  mutate(agegp.lab = fct_recode(agegp, 
                                !!!setNames(paste(agegp.labels.used), names(agegp.labels.used)))) %>%
  mutate(sex.lab = ifelse(sex==1, 'Males', 'Females')) %>%
  mutate(qty.lab = fct_recode(quantity,
                              !!!qty.names)) %>%
  ggplot(.) +
    geom_path(aes(x=agegp.lab, y=asdr, color=qty.lab, group=interaction(qty.lab, sex))) +
    facet_grid(sex.lab ~ .) +
    xlab('') +
    ylab('Simulated death rate') +
    guides(color=guide_legend(
                              #title='Quantity',
                              title='',
                              position='bottom',
                              nrow=2)) +
    theme_minimal() +
    theme(legend.position='bottom')

all.truth.plot

pdf(here('sim-out', "sim-all-truths-m.pdf"), width=6, height=4)
print(all.truth.plot)
dev.off()
```

## Plot the age-sex distribution of the (simulated) universe of siblings

```{r}
qty.names <- c('Total population'='truth',
               'Visible population'='visible_truth',
               'Visible population (inc. resp.)' = 'visible_truth_withr',
               'Invisible population'='invisible_truth')

all.truth.exp.plot <-  all.truths.forplots %>%
  separate(cell, into=c('agegp', 'sex'), sep='X') %>%
  mutate(agegp=str_extract(agegp, '\\d'),
         sex=str_extract(sex, '\\d')) %>%
  mutate(agegp.lab = fct_recode(agegp, 
                                !!!setNames(paste(agegp.labels.used), names(agegp.labels.used)))) %>%
  mutate(sex.lab = ifelse(sex==1, 'Males', 'Females')) %>%
  mutate(qty.lab = fct_recode(quantity,
                              !!!qty.names)) %>%
  ggplot(.) +
    geom_path(aes(x=agegp.lab, y=N.cell, color=qty.lab, group=interaction(qty.lab, sex))) +
    facet_grid(sex.lab ~ .) +
    xlab('') +
    ylab('Simulated population size') +
    guides(color=guide_legend(
                              title='',
                              #title='Quantity',
                              position='bottom',
                              nrow=2)) +
    theme_minimal() +
    theme(legend.position='bottom')

all.truth.exp.plot

pdf(here('sim-out', "sim-all-truths-n.pdf"), width=6, height=4)
print(all.truth.exp.plot)
dev.off()
```

### Summarize the estimates

```{r}
res.summ <- svy.sims.forplots %>%
  filter(! cell %in% c('agegp1Xsex1', 'agegp1Xsex2', 'agegp9Xsex1', 'agegp9Xsex2')) %>%
  group_by(f, cell, estimator, uber.tau.D, uber.tau.N) %>%
  summarize(asdr = mean(asdr.hat),
            asdr.ci.low = quantile(asdr.hat, .025),
            asdr.ci.high = quantile(asdr.hat, .975))

res.summ
```


# Plot the simulation results

## Plots that demonstrate the sensitivity frameworks are correct

Construct a dataset that compares the estimands, adjusted estimands, and true visible death rates.
If the sensitivity frameworks are correct, then the adjusted estimands should be equal to the true
visible death rates.

```{r}
# results to plot for estimators without respondent included
sens.check <- all.estimands.forplots %>%
          select(cell, asdr, uber.tau.N, uber.tau.D, quantity=estimand) %>%
          filter(quantity %in% c('estimand_agg', 'estimand_agg_adj',
                                 'estimand_ind', 'estimand_ind_adj')) %>%
          separate(quantity, 
                   into=c('foo', 'estimator', 'adjusted'), 
                   sep="\\_", fill='right')   %>%
          select(-foo) %>%
          mutate(adjusted = case_when(is.na(adjusted) ~ 'unadjusted',
                                      TRUE ~ 'adjusted')) %>%
          ungroup() %>%
          left_join(all.truths.forplots %>%
                      filter(quantity %in% c('visible_truth')) %>%
                      select(cell, true_visible_asdr=asdr),
                    by='cell') %>%
          ## split cell into readable age/sex groups
          separate(cell, into=c('agegp', 'sex'), sep='X') %>%
          mutate(agegp = str_extract(agegp, '\\d'),
                 sex = str_extract(sex, '\\d')) %>%
          mutate(agegp.lab = fct_recode(agegp, 
                                    !!!setNames(paste(agegp.labels.used), 
                                                names(agegp.labels.used)))) %>%
          mutate(sex.lab = ifelse(sex==1, 'M', 'F'))

# results to plot for estimators with respondents included
sens.check.withR <- all.estimands.forplots %>%
          select(cell, asdr, uber.tau.N, uber.tau.D, quantity=estimand) %>%
          ## TODO !!!
          ## need to calculate and use estimand_agg_withR_adj and estimand_ind_withR_adj
          filter(quantity %in% c('estimand_agg_withR', 'estimand_agg_withR_adj',
                                 'estimand_ind_withR', 'estimand_ind_withR_adj'))  %>%
          mutate(adjusted = case_when(quantity=='estimand_agg_withR_adj' ~ 'adjusted',
                                      quantity=='estimand_ind_withR_adj' ~ 'adjusted',
                                      TRUE ~ 'unadjusted'),
                 estimator = case_when(quantity=='estimand_agg_withR' ~ 'agg_withR',
                                       quantity=='estimand_agg_withR_adj' ~ 'agg_withR',
                                       TRUE ~ 'ind_withR'))  %>%
          ungroup() %>%
          ## split cell into readable age/sex groups
          left_join(all.truths.forplots %>%
                      filter(quantity %in% c('visible_truth_withr')) %>%
                      select(cell, true_visible_asdr=asdr),
                    by='cell') %>%
          separate(cell, into=c('agegp', 'sex'), sep='X') %>%
          mutate(agegp = str_extract(agegp, '\\d'),
                 sex = str_extract(sex, '\\d')) %>%
          mutate(agegp.lab = fct_recode(agegp, 
                                    !!!setNames(paste(agegp.labels.used), names(agegp.labels.used)))) %>%
          mutate(sex.lab = ifelse(sex==1, 'M', 'F')) 
         
```

These will be helpful for labeling the next few plots...

```{r}
appender_tau_N <- function(string) {
    TeX(paste("$\\tau_N = $", string))
}
appender_tau_D <- function(string) {
    TeX(paste("$\\tau_D = $", string))
}
```

### Aggregate estimator - not including respondent

```{r}

## check aggregate estimator without respondent included
agg.check.plot <-  ggplot(sens.check %>% 
                            filter(estimator=='agg')) +
  geom_abline(intercept=0, slope=1, color='grey') +
  geom_point(aes(x=true_visible_asdr, 
                 y=asdr, 
                 color=sex.lab, 
                 shape=adjusted)) +
  facet_grid(uber.tau.N ~ uber.tau.D, 
             labeller=labeller(uber.tau.D = as_labeller(appender_tau_D, default=label_parsed),
                               uber.tau.N = as_labeller(appender_tau_N, default=label_parsed))) +
  xlab("True visible death rate") +
  ylab("Death rate estimand") +
  scale_color_discrete(name='') +
  scale_shape_discrete(name='') +
  ggtitle('Aggregate visibility estimator') +
  theme_minimal()

pdf(here('sim-out', "agg-check-plot.pdf"), width=10, height=10)
print(agg.check.plot)
dev.off()

agg.check.plot
```

### Individual estimator - not including respondent

```{r}
## check individual estimator without respondent included
ind.check.plot <-  ggplot(sens.check %>% 
                            filter(estimator=='ind')) +
  geom_abline(intercept=0, slope=1, color='grey') +
  geom_point(aes(x=true_visible_asdr, 
                y=asdr, 
                color=sex.lab, 
                shape=adjusted)) +
  facet_grid(uber.tau.N ~ uber.tau.D, 
             labeller=labeller(uber.tau.D = as_labeller(appender_tau_D, default=label_parsed),
                               uber.tau.N = as_labeller(appender_tau_N, default=label_parsed))) +
  xlab("True visible death rate") +
  ylab("Death rate estimand") +
  scale_color_discrete(name='') +
  scale_shape_discrete(name='') +
  ggtitle('Individual visibility estimator') +
  theme_minimal()

pdf(here('sim-out', "ind-check-plot.pdf"), width=10, height=10)
print(ind.check.plot)
dev.off()

ind.check.plot

```

### Aggregate estimator - including respondent

```{r}
## check aggregate estimator without respondent included
agg.withR.check.plot <-  ggplot(sens.check.withR %>% 
                            filter(estimator=='agg_withR')) +
  geom_abline(intercept=0, slope=1, color='grey') +
  geom_point(aes(x=true_visible_asdr, 
                 y=asdr, 
                 color=sex.lab, 
                 shape=adjusted)) +
  facet_grid(uber.tau.N ~ uber.tau.D, 
             labeller=labeller(uber.tau.D = as_labeller(appender_tau_D, default=label_parsed),
                               uber.tau.N = as_labeller(appender_tau_N, default=label_parsed))) +
  xlab("True visible death rate\n(Including respondent)") +
  ylab("Death rate estimand") +
  scale_color_discrete(name='') +
  scale_shape_discrete(name='') +
  ggtitle('Aggregate visibility estimator (respondent included)') +
  theme_minimal()

pdf(here('sim-out', "agg-withR-check-plot.pdf"), width=10, height=10)
print(agg.withR.check.plot)
dev.off()

agg.withR.check.plot
```

### Individual estimator - including respondent

```{r}
## check individual estimator without respondent included
ind.withR.check.plot <-  ggplot(sens.check.withR %>% 
                            filter(estimator=='ind_withR')) +
  geom_abline(intercept=0, slope=1, color='grey') +
  geom_point(aes(x=true_visible_asdr, 
                y=asdr, 
                color=sex.lab, 
                shape=adjusted)) +
  facet_grid(uber.tau.N ~ uber.tau.D, 
             labeller=labeller(uber.tau.D = as_labeller(appender_tau_D, default=label_parsed),
                               uber.tau.N = as_labeller(appender_tau_N, default=label_parsed))) +
  xlab("True visible death rate\n(including respondent)") +
  ylab("Death rate estimand") +
  scale_color_discrete(name='') +
  scale_shape_discrete(name='') +
  ggtitle('Individual visibility estimator (respondent included)') +
  theme_minimal()
pdf(here('sim-out', "ind-withR-check-plot.pdf"), width=10, height=10)

print(ind.withR.check.plot)
dev.off()

ind.withR.check.plot

```
## Plots that summarize the error

Summarize the MSE wrt true visible death rate (excl. respondent) for all estimators

```{r}
dodge.width <- .75

est.names <- c('Aggregate'='agg',
               'Aggregate (with resp.)'='agg_withR',
               'Individual'='ind',
               'Individual (with resp.)'='ind_withR')

plots.mse.summ <- svy.sims.forplots %>%
  split(interaction(.$f, .$uber.tau.D, .$uber.tau.N)) %>%
  
  map(function(df) {
    
    mse.summ <- df %>%
      left_join(visible_truth.forplots %>% 
                  select(cell, true.vis.asdr=asdr),
                by=c('cell')) %>%
      group_by(estimator, cell) %>%
      dplyr::summarise(mse = mean((1000*asdr.hat - 1000*true.vis.asdr)^2),
                       var = var(1000*asdr.hat),
                       bias2 = (1000*mean(true.vis.asdr-asdr.hat))^2)  %>%
      separate(cell, into=c('agegp', 'sex'), sep='X') %>%
      mutate(agegp = str_extract(agegp, '\\d'),
             sex = str_extract(sex, '\\d'))
    
    toplot <- mse.summ %>%
      ungroup() %>%
      dplyr::mutate(agegp.lab = fct_recode(agegp, 
                                           !!!setNames(paste(agegp.labels.used), 
                                                       names(agegp.labels.used)))) %>%
      dplyr::mutate(sex.lab = ifelse(sex==1, 'M', 'F')) %>%
      dplyr::mutate(est.lab = as_factor(estimator),
             est.lab = fct_recode(est.lab,
                                  !!!est.names))
        
    mse.plot <- ggplot(toplot) +
      #geom_bar(aes(x=agegp, y=mse, fill=estimator), stat='identity', position=position_dodge()) +
      geom_point(aes(x=agegp.lab, y=mse, color=est.lab, shape='mse'), 
                 stat='identity', 
                 position=position_dodge(width=dodge.width)) +
      geom_point(aes(x=agegp.lab, y=var, color=est.lab, shape='var'), 
                 stat='identity', 
                 position=position_dodge(width=dodge.width)) +
      geom_point(aes(x=agegp.lab, y=bias2, color=est.lab, shape='bias2'), 
                 stat='identity', 
                 position=position_dodge(width=dodge.width)) +
      facet_grid(sex.lab ~ .) +
      scale_shape_manual(name='quantity',
                         values=c(1,3,4)) +
      xlab('') +
      ylab('Mean square error\n(w.r.t. visible truth excl. R)') +
      ggtitle(TeX(glue(
        "Sampling fraction: {df$f[1]}, $\\tau_N$: {df$uber.tau.N[1]}; $\\tau_D$: {df$uber.tau.D[1]}",
        ))) +
      guides(color=guide_legend(title='Estimator'),
             shape=guide_legend(title='Quantity')) +
      theme_minimal() +
      theme(legend.position='bottom',
            legend.box='vertical')
    
    return(mse.plot)
  })

pdf(here('sim-out', "prelim-sim-mse-results.pdf"), width=7, height=4)
print(plots.mse.summ)
dev.off()

pdf(here('sim-out', "sim-mse-results-f3.pdf"), width=7, height=7)
print(plots.mse.summ[[3]])
dev.off()

plots.mse.summ
```

Summarize the MSE wrt true visible death rate (excl. respondent) for estimators that exclude respondents only

```{r}
dodge.width <- .5

est.names <- c('Aggregate'='agg',
               'Aggregate (with resp.)'='agg_withR',
               'Individual'='ind',
               'Individual (with resp.)'='ind_withR')

## scaling factor by which death rates get multiplied
## (sometimes it's nice to have death rates per 1,000, etc)
Kasdr <- 1000

plots.mse.exclr <- svy.sims.forplots %>%
  split(interaction(.$uber.tau.D, .$uber.tau.N, .$f)) %>%
  map(function(df) {
    
    mse.summ <- df %>%
      left_join(visible_truth.forplots %>% select(cell, true.vis.asdr=asdr),
                by=c('cell')) %>%
      group_by(estimator, cell) %>%
      dplyr::summarise(
                relmse = (mean((Kasdr*asdr.hat - Kasdr*true.vis.asdr)^2)/(Kasdr*true.vis.asdr[1])),
                relvar = var(Kasdr*asdr.hat)/(Kasdr*true.vis.asdr[1]),
                relbias2 = ((Kasdr*mean(true.vis.asdr-asdr.hat))^2)/(Kasdr*true.vis.asdr[1])) %>%
      separate(cell, into=c('agegp', 'sex'), sep='X') %>%
      dplyr::mutate(agegp = str_extract(agegp, '\\d'),
             sex = str_extract(sex, '\\d')) %>%
      dplyr::mutate(agegp.lab = fct_recode(agegp, !!!setNames(paste(agegp.labels.used), 
                                                              names(agegp.labels.used)))) %>%
      dplyr::mutate(sex.lab = ifelse(sex==1, 'm', 'f'))

      toplot <- mse.summ %>%
        ungroup() %>%
        dplyr::mutate(agegp.lab = fct_recode(agegp, 
                                  !!!setNames(paste(agegp.labels.used), 
                                              names(agegp.labels.used)))) %>%
        dplyr::mutate(sex.lab = ifelse(sex==1, 'M', 'F')) %>%
        dplyr::mutate(est.lab = as_factor(estimator),
               est.lab = fct_recode(est.lab,
                                    !!!est.names)) 
    
    mse.plot.exclr <- ggplot(toplot %>% 
                               ungroup() %>%
                               filter(estimator %in% c('agg', 'ind'))  %>% 
                               mutate(est.lab=fct_drop(est.lab))) +
      #geom_bar(aes(x=agegp, y=mse, fill=estimator), stat='identity', position=position_dodge()) +
      geom_point(aes(x=agegp.lab, y=relmse, color=est.lab, shape='mse'), 
                 stat='identity', 
                 position=position_dodge(width=dodge.width)) +
      geom_point(aes(x=agegp.lab, y=relvar, color=est.lab, shape='var'), 
                 stat='identity', 
                 position=position_dodge(width=dodge.width)) +
      geom_point(aes(x=agegp.lab, y=relbias2, color=est.lab, shape='bias2'), 
                 stat='identity', 
                 position=position_dodge(width=dodge.width)) +
      facet_grid(sex.lab ~ .) +
      scale_shape_manual(name='quantity',
                         values=c(1,3,4)) +
      ylab('Relative mean square error\n(w.r.t. visible truth excl. R)') +
      xlab('') +
      ggtitle(paste0('Sampling fraction: ', df$f[1], 
                     '; tau.D: ', df$uber.tau.D[1], 
                     '; tau.N: ', df$uber.tau.N[1])) +
      theme_minimal() +
      guides(color=guide_legend(title='Estimator'),
             shape=guide_legend(title='Quantity')) +
      theme(legend.position='bottom',
            legend.box='vertical')
          
    return(mse.plot.exclr)
  })

pdf(here('sim-out', "prelim-sim-mse-exclr-results.pdf"), width=7, height=4)
print(plots.mse.exclr)
dev.off()

pdf(here('sim-out', "sim-mse-results-withoutresp-f05-imperfectreporting.pdf"), width=7, height=7)
print(plots.mse.exclr[[3]])
dev.off()

pdf(here('sim-out', "sim-mse-results-withoutresp-f05-perfectreporting.pdf"), width=7, height=7)
print(plots.mse.exclr[[4]])
dev.off()

plots.mse.exclr
```



```{r}
toc()
```



























